On the Duality of Frequentist Point Estimates and Bayesian Shrinkage Priors: An Extension based on the Triple-Gamma-Prior

### 2024-10-06 01_latex/MScThesis_Econ.tex
Pre-Fix Up Commit

\begin{abstract}
This thesis introduces the Triple-Gamma Regularization (TGR), a novel non-convex penalty term inspired by the Triple-Gamma prior, which provides increased flexibility in handling under- and overfitting in high-dimensional regression problems. The proposed method offers a unifying framework that bridges convex penalties like LASSO and non-convex penalties such as Gaussian and Arctan, allowing for fine-tuning of the regularization process through adjustable hyperparameters. First, a detailed theoretical derivation of the TGR is provided, highlighting its properties and advantages over existing regularization methods. Then, a comprehensive simulation study is conducted to evaluate the performance of TGR under various high-dimensional scenarios. The results indicate that TGR can outperform traditional regularization techniques in several settings, particularly in terms of minimizing mean squared error. This thesis also addresses the computational challenges associated with non-convex optimization problems and explores potential solutions through gradient-based optimization techniques. These findings contribute to the growing body of literature on regularization methods and offer practical insights for future applications in sparse modeling and high-dimensional data analysis.
\end{abstract}

@@

This completion oft this thesis would not have been possible without so many amazing people in my life who supported me throughout the process of writing this master's thesis and the master program as a whole. I want to thank my mum, Daniela, and my dad, Werner, who always picked me up when I lost my motivation to continue and who convinced me to keep going time after time. I want to thank my brother, Tobias, who always readily listened to me complaining for hours and hours on end about things he probably barely understood and yet always agreed with me about my computational issues while doing the simulations. And most of all I want to thank my advisor, Peter, who always supported me from the day I randomly dropped an e-mail into his e-mail inbox about becoming his tutor in econometrics, although not having met once during the distance learning phase of the COVID pandemic, up until today. From working as a tutor in his team, writing a bachelor thesis, and to the support during this master thesis now. Even though sometimes living 12h time differences apart, one in Boston the other in Manila, he always answered any questions I had and supported me in whatever way he could and I want to sincerely thank him for that!

@@

A prior distribution which might have these suitable properties is the unifying shrinkage prior developed by \textcite{TGP2020} called \textit{Triple-Gamma-Prior}. Its unifying properties and the fact that a closed-form solution for the marginal prior exists might lead to favourable properties in the realm of frequentist statistics. Before that it is necessary to have a closer look at what the exact problem is that Bayesian shrinkage priors and regularization approaches are trying to solve. An overview of this along with a theoretical basis of the necessary concepts will be presented in section 2.\\

@@

The first of these two goals - \textit{inference} - aims at understanding the relationship that may or may not exist between the input variables $\mathbf{X_i}$ and the output $\mathbf{Y}$. Especially applied sciences like Economics, Psychology and Medicine often try to find a (causal) relationship within their theoretical framework to evaluate a policy, a medication or a new form of therapy. Linear models for example often provide a simple and straightforward framework which provide the scientist with interpretable effects. \textit{Prediction} on the other hand aims at forecasting the output variable $Y$ as accurately as possible and using every bit of information that is available, but disregards interpretability \parencite[21]{ESL2009}. These goals can be summarized by viewing it as a decision between \textit{prediction accuracy} and \textit{model interpretability}, which has thoroughly been explained by \textcite{ESL2009}.\\

@@

In supervised learning, a model is fitted on an observed dataset where both inputs and outputs are given in pairs. Using this data a statistical model aims to find a functional form which maps these inputs into the output space and thus describes a relationship between those two spaces. Naturally, one would assume to find a functional form which manages to describe the relationship as closely as possible by minimizing a chosen loss function. Yet, this often ignores a problem called "overfitting" which can loosely be described as using "too much information" from the training dataset. As described by \textcite{Ying2019}, this problem arises from the fact "that [an] over-fitted model has difficulty coping with pieces of the information in the testing set, which may be different from those in the training set". A more formal definition of overfitting has been defined by \textcite{Mitchell1997}:\\

@@

In general, a linear regression model fitted with least squares on a sufficient amount of data points ($n \gg p$) will produce estimates which both have low bias and low variance and thus tends to perform well out of sample. However, problems arise with this approach arise when the sample size decreases, because the variance in the estimates will increase. The main idea behind this can be easily be visualized by trying to fit a polynomial curve on a two dimensional space and then altering the available sample for polynomials of different order. 
\begin{figure}
\centering
\includegraphics[scale=0.5]{PolynomialFit_Plot.png}
\caption{1st and 8th order polynomial fit to data \\ (\textcolor{green1}{Green} \& \textcolor{blue1}{Blue} Lines fitted with red data point; \textcolor{orange1}{Orange} \& \textcolor{red1}{Red} Lines fitted without).}
\label{fig:polyfit}
\end{figure}

@@

Let $\mathcal{D}$ be a training data set with $n \in \mathbb{N}$ observations, where every data point consists of a target variable $y_i \in \mathbb{R}$ along with a number of corresponding inputs $x_i \in \mathbb{R}^p$. Given a linear function $f_\beta(x_i)$ of the inputs parametrized by coefficients $\beta \in \boldsymbol{\beta}$, $L(y_i f_\beta(x_i))$ is the \textit{Loss} function measuring the discrepancy between the actual target $y_i$ and the output of the linear function $f_\beta(x_i)$. According to \textcite{Vapnik1991}, the \textit{Empirical Risk Functional} is then 
\[
R_{emp}(\beta) = \frac{1}{n} \sum_{i=1}^n L(y_i, f_\beta(x_i)).
\]
\textit{Regularization} in this thesis' context refers then to adding some penalty function $J(\beta)$ dependent on the set of parameters $\beta$, multiplied by some weighting parameter $\lambda$, to the empirical risk functional $R_{emp}(\beta)$. Thus, $R_{reg} = R_{emp}(\beta) + \lambda\cdot J(\beta)$. This results in the overall optimization problem
\begin{align}
&\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{R_{reg}(\beta)\right\}\nonumber\\
&=\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{R_{emp}(\beta) + \lambda\cdot J(\beta)\right\}\nonumber\\
&=\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{\sum_{i=1}^n L(y_i, f_\beta(x_i)) + \lambda\cdot J(\beta)\right\}.\label{eq:genreg}
\end{align}

It is important to note here that $J(\beta)$ is only a function of the coefficients $\beta$ and neither the targets $y_i$ nor the inputs $x_i$. Therefore, it only affects the generalization error of the model, not the training error given by the empirical risk functional $R_{emp}(\beta)$.

@@

The concept of a penalized regression has been around for quite some time and been studied widely in various fields of scientific research. Arguably, this methodological approach to penalized regression started with the publication of two pieces of literature published by Arthur Hoerl and Robert Kennard in 1970 \parencite{HoerlKennard1970a, HoerlKennard1970b}. With these two papers, the authors introduced the widely known \textit{Ridge Regression}, which was derived from the previously known concept of Ridge analysis. At its core, the authors were trying to tackle the problem of high variances of the regression coefficients in high-dimensional problem settings. This shrinkage estimator, which uses the squared coefficient as a penalty term, "attempt[s] to shrink the coefficients to
reduce these variances, while adding some bias." \parencite{Hoerl2020} This closely resembles the previously discussed issue of the \textit{Bias-Variance-Tradeoff}, which has been discussed in the \textit{Under- and Overfitting} chapter in section \ref{sec:theorysuper} (Roger W. Hoerl, Arthur Hoerl's son, published a historical overview of the development of the concept of \textit{Ridge Regression} in 2020 \parencite{Hoerl2020}). The closed from solution of the Ridge estimator is given by
$$\hat{\beta}_{Ridge} = \left(X^TX+\lambda I\right)^{-1}X^Ty,$$
which adjusts the OLS estimator by shifting the main diagonal entries of the design matrix by $\lambda$ ($\lambda \geq 0$). It can be shown that this closed form estimator is equivalent to a Lagrangian problem of the following form \parencite{VanWieringen2015}:
$$\hat{\beta}_{Ridge}(\lambda) = \underset{\beta}{\argmin}\left\{\norm{y-X\beta}_2^2 + \lambda\norm{\beta}_2^2\right\}.$$
This resembles an example of the above defined regularization framework with squared residual loss and a penalty term of the form $\norm{\beta}_2^2$, which only depends on the parameter $\beta$. In case of $\lambda$ being equal to zero, this reduces to the \textit{maximum likelihood (ML) estimator}.

@@

$$\norm{\beta}_p =  \left(\sum_{i=1}^p|\beta_i|^p\right)^{1/p}.$$

Work published by researchers in the nineties, like the previously mentioned \textcite{FrankFriedman1993} or \textcite{Fu1998}, as well as more recent literature like \textcite{WangEtAl2020} have repeatedly shown there is no go-to-method to tackle regularization problems, as the effectiveness of a specific approach highly depends on the data siuation at hand. Due to this particular situation in the literature, several other methods have been proposed in the recent years and decades. An approach combining the \textit{Ridge} and \textit{LASSO} methods is called \textit{Elastic Net} regularization and has been developed by \textcite{ZouHastie2005}. The authors there elaborate on some of the shortcomings of the LASSO method. For example, in a special case where there are more predictors $p$ than data points $n$ ($p > n$), LASSO only selects up to $n$ variables due to the nature of the convex optimization problem. Should several of the included variables be highly \textit{pairwise} correlated with each other, \textit{LASSO} tends to only select on of these variables. In its core, \textit{Elastic Net Regularization} linearly combines the penalty terms of \textit{Ridge} and \textit{LASSO} regularization, yielding a loss function of the form:\\
$$J(\beta) = \lambda_1 \norm{\beta}_2^2 + \lambda_2 \norm{\beta}_1.$$
The authors have	 shown that, especially when it comes to encouraging the aforementioned grouping effects, \textit{Elastic Net} tends to perform better than the \textit{LASSO}.\\

@@

Seeing this vast array of different pieces of literature immediately raises the question on advantages and disadvantages of specific methods and why so many have established itself in this particular field of study. Keeping in mind that the effectiveness of a method still highly depends on the data situation at hand, it is still important to distinguish methods based on its mathematical properties. \textcite{JohnVettamWu2022} have distinguished methods based on two broad criteria: (1) How does a penalty behave in small neighbourhoods around zero? (Is it smooth or singular at origin?) and (2) Is the underlying function convex or non-convex? To better incorporate the proposed concept of this thesis into this body of literature, I am proposing a third property to distinguish penalty terms: (3) How does the function behave in the limits? (Does it converge or diverge?) Such a classification tree can be found in figure \ref{fig:Classification}. \\

\begin{figure}
\begin{center}
\begin{tikzpicture}[
    level 1/.style={sibling distance=5.2cm},
    level 2/.style={sibling distance=2.6cm},
    level 3/.style={sibling distance=2.6cm, level distance=2.5cm},
    every node/.style={draw, rectangle, align=center}
]
\node {Classification of \\ Penalty Terms}
    child { node {Behaviour at Zero}
        child { node {Singularity} }
        child { node {Smooth at \\ Origin}}
    }
    child { node {(Non-)Convexity}
        child { node {Convex}}
        child { node {Non-Convex}}
    }
    child { node {Behaviour at Limits}
        child { node {Divergent}}
        child { node {Convergent}}
    };
\end{tikzpicture}  
\end{center}
\caption{A possible classification of regularization methods based on the mathematical properties of the penalty function} \label{fig:Classification}
\end{figure}
As already mentioned earlier in section \ref{sec:theorysuper}, a penalty function is a function dependent on the parameter of the model and it does not depend on the data at hand. Thus, changes in your data set does not effect the penalty directly, but only the overall optimization problem through the empirical risk functional $R_{emp}$. The following table \ref{tab:ClassificationOfPenalties} classifies some of the existing concepts based on the three criteria mentioned earlier. A visualisation of them can be found in figure \ref{fig:DifferentPenalties} in chapter \ref{subsec:comparepenalties}.\\

@@

LASSO 	& \textcite{Tibishirani1996}	& Singular	&	C 	& Div.\\
Ridge 	& \textcite{HoerlKennard1970a}	& Smooth		& C			& Div.\\
Gaussian & \textcite{JohnVettamWu2022}	& Smooth		& NC 	& Conv.\\
Ar(c)tan 	& \textcite{WangZhu2016}	& Singular 	& NC 	& Conv. \\
Elastic-Net & \textcite{ZouHastie2005} & Singular & C &\\
Bridge & \textcite{FrankFriedman1993}& Both &&\\
MCAP & \textcite{Zhang2010}& Singular & NC &\\
SCAD & \textcite{FanLi2001} & Singular & NC &\\
Laplace & \textcite{TrzaskoManduca2009} & NC &&\\
Triple-Gamma & - & \makecell{Depending\\ on \\ parameters} & NC & Div. \\

@@

The main motivation for this thesis stems from a striking duality that exists between Bayesian shrinkage priors (discussed in chapter \ref{sec:BayesianShrinkagePriors}) and the frequentist approach using regularization. Both approaches aim at tackling the issue of model complexity and overfitting by making it necessary for the data to be more convincing that the value of an estimate is statistically significant different from zero. As mentioned, Bayesian statistics use specific prior distributions with a usually a lot of mass around zero and heavy tails, whereas frequentist statisticians usually alter their optimization problems incorporating penalty terms into into their loss functions. On a first glance, these two approaches have no immediate mathematical connection, however this turns out to be wrong.\\

@@

It turns out, as can be seen in \textcite[248--250]{ESL2009} and \textcite{VanWieringen2015} among others, that when choosing certain prior distributions and using standard assumptions in Bayesian modelling (ie the individual parameters of the parameter vector are independent a priori and the errors of the standard linear model are drawn from a normal distribution), that the mode of the posterior distributions correspond to the point estimates of a regularization approach in a frequentist setting. As this idea is essential to the entire thesis, a shortened derivation of this duality for the case of \textit{Ridge} regression will now be shown.\\

Assuming that a response variable $y_i$ follows a linear model with $p$ variables and errors $\epsilon_i \sim \mathcal{N}(0, \sigma^2)$, then the likelihood of the data is given by
$$\mathcal{L}(y|X,\beta)=\frac{1}{(2\pi\sigma^2)^{n/2}}\exp\left(-\frac{1}{2\sigma^2}\sum_{i=1}^n(y_i-x_i^T\beta)^2\right).$$

Under the previously mentioned assumption that $p(\beta) = \prod_{i=1}^k \beta_i$ and by choosing a Gaussian distribution with mean zero and variance $\tau^2$, we gain the prior distribution
$$p(\beta) = \frac{1}{(2\pi\tau^2)^{p/2}}\exp\left(-\frac{1}{2\tau^2}\norm{\beta}_2^2\right).$$

Putting both together using Bayes' theorem and taking the $\log$ of the distribution yields
$$\log(p(\beta|y,X)) \propto -\frac{1}{2\sigma^2}\sum_{i=1}^n (y_i-x_i^T\beta)^2  - \frac{1}{2\tau^2}\norm{\beta}_2^2.$$

By viewing $\frac{1}{2\tau^2}$ as the weighting parameter $\lambda$, which controls the strength of the regularization. The \textit{maximum-a-priori} estimate $\hat{\beta}_{MAP}$, which can be viewed as trying to find the $\underset{\beta}{\argmax}$ of the $\log$-posterior, can be used to construct the following optimization problem:

$$\hat{\beta}_{MAP} = \underset{\beta}{\argmin}\left\{\frac{1}{2\sigma^2}\sum_{i=1}^n \epsilon_i^2  + \frac{1}{2\tau^2}\norm{\beta}_2^2\right\}.$$

This optimization problem constitutes the connection between the Bayesian and the frequentist approach as it can be interpreted as (1) finding the mode of the posterior distribution when using a Gaussian prior distribution and (2) finding the parameters which minimize the loss when using a Ridge regression with $\lambda = 1/(2\tau^2)$. The same connection exists for the case of \textit{LASSO} regression which can be derived using the same procedure and a double-exponential - also called \textit{Laplace} - distribution with mean zero and some scale parameter $\tau$ as the prior for the parameter vector $\beta$.

@@

As already mentioned above, shrinkage priors are a Bayesian approach to dealing with statistical problems in high-dimensional settings. The prior is generally constructed for variance selection in sparse state space and TVP models and is named the Triple-Gamma-Prior as its representation is given by three gamma distributions. A key advantage of this prior is its unifying property. As can be seen in table 1 in \textcite{TGP2020}, special cases of the prior, depending on certain values in its hyperparameters, are the normal-exponential-Gamma, the horseshoe and the double gamma prior. In addition, the Bayesian LASSO can be seen as a special case of the triple-gamma-prior. Another advantage is its function as a bridge between spike-and-slab and continuous shrinkage prior as the hyperparameters can be used both for steering the shrinkage effect on noise while also preventing overshrinking the signal. Computationally the shrinkage prior performs better than commonly used spike-and-slab priors as it is a continuous shrinkage prior which can more easily be implemented using Markov-Chain-Monte-Carlo (MCMC) methods.\\

@@

%----- Model Setup and Derivation
\newpage
\section{Model Setup and Derivation}\label{sec:modelderivation}
Coming to the theoretical framework of the \textit{triple-gamma-regularization}, let's assume we have a response variable $y$ and $p$ predictors along with $n$ data points. More formally, let $y=[y_1  \quad y_2 \cdots y_n]^T$ and $x_i = [x_{i1} \quad x_{i2} \cdots x_{ip}]^T$ with $\forall i\in \{1,...,n\}: y_i, x_{ij} \in \mathbb{R}$. Here, $x_i$ is the $i$-th predictor, thus resulting in the design matrix $X = [x_1 \quad x_2 \cdots x_p]$.\\

>>>>

\begin{abstract}
This thesis introduces the Triple-Gamma Regularization (TGR), a novel non-convex penalty term inspired by the Triple-Gamma prior, which provides increased flexibility in handling under- and overfitting in high-dimensional regression problems. The proposed method offers a unifying framework that bridges convex penalties like LASSO and non-convex penalties showing similar characteristics as Gaussian and Arctan penalites, allowing for fine-tuning of the regularization process through adjustable hyperparameters. First, a detailed theoretical derivation of the TGR is provided alongside an analysis of its properties and advantages over existing regularization methods. Then, a comprehensive simulation study is conducted to evaluate the performance of TGR under various high-dimensional scenarios. The results indicate that TGR can outperform traditional regularization techniques in several settings, specifically in terms of minimizing the mean squared error on validation sets. This thesis also addresses the computational challenges associated with non-convex optimization problems and explores potential solutions through gradient-based optimization techniques. These findings contribute to the growing body of literature on regularization methods and offer practical insights for future applications in sparse modelling and high-dimensional data analysis.
\end{abstract}

@@

The completion of this thesis would not have been possible without so many amazing people in my life who supported me throughout the process of writing this master's thesis and the master program as a whole. I want to thank my mum, Daniela, and my dad, Werner, who always picked me up when I lost my motivation to continue and who convinced me to keep going time after time. I want to thank my brother, Tobias, who always readily listened to me complaining for hours and hours on end about things he probably barely understood and yet always agreed with me about my computational issues while doing the simulations. And most of all I want to thank my advisor, Peter, who always supported me from the day I randomly dropped an e-mail into his e-mail inbox about becoming his tutor in econometrics, although not having met me once during the distance learning phase of the COVID pandemic, up until today. From working as a tutor in his team, writing a bachelor thesis, and to the support during this master thesis now. Even though sometimes living 12h time differences apart, one in Boston and the other in Manila, he always answered any questions I had and supported me in whatever way he could and I want to sincerely thank him for that!

@@

A prior distribution which might have these suitable properties is the unifying shrinkage prior developed by \textcite{TGP2020} called \textit{Triple-Gamma-Prior}. Its unifying properties and the fact that a closed-form solution for the marginal prior exists might lead to favourable properties in the realm of frequentist statistics. Before that, it is necessary to have a closer look at what the exact problem is that Bayesian shrinkage priors and regularization approaches are trying to solve. An overview of this along with a theoretical basis of the necessary concepts will be presented in section 2.\\

@@

The first of these two goals - \textit{inference} - aims at understanding the relationship that may or may not exist between the input variables $\mathbf{X_p}$ and the output $\mathbf{Y}$. Especially applied sciences like economics, psychology and medicine often try to find a (causal) relationship within their theoretical framework to evaluate a policy, a medication or a new form of therapy. Linear models for example often provide a simple and straightforward framework which provide the scientist with interpretable effects. \textit{Prediction} on the other hand aims at forecasting the output variable $\mathbf{Y}$ as accurately as possible and using every bit of information that is available, but disregards interpretability \parencite[21]{ESL2009}. These goals can be summarized by viewing it as a decision between \textit{prediction accuracy} and \textit{model interpretability}, which has thoroughly been explained by \textcite{ESL2009}.\\

@@

In supervised learning, a model is fitted on an observed dataset where both inputs and outputs are given in pairs. Using this data, a statistical model aims to find a functional form which maps these inputs into the output space and thus describes a relationship between those two spaces. Naturally, one would assume to find a functional form which manages to describe the relationship as closely as possible by minimizing a chosen loss function. Yet, this often ignores a problem called "overfitting" which can loosely be described as using "too much information" from the training dataset. As described by \textcite{Ying2019}, this problem arises from the fact "that [an] over-fitted model has difficulty coping with pieces of the information in the testing set, which may be different from those in the training set". A more formal definition of overfitting has been defined by \textcite{Mitchell1997}:\\

@@

In general, a linear regression model fitted with least squares on a sufficient amount of data points ($n \gg p$) will produce estimates which both have low bias and low variance and thus tends to perform well out-of-sample. However, problems arise with this approach arise when the sample size decreases, because the variance in the estimates will increase. The main idea behind this can be easily be visualized by trying to fit a polynomial curve on a two dimensional space with only a few data points and then altering the available sample for polynomials of different order. 
\begin{figure}
\centering
\includegraphics[scale=0.5]{PolynomialFit_Plot.png}
\caption{1st and 8th order polynomial fit to data \\ (\textcolor{green1}{Green} \& \textcolor{blue1}{Blue} lines fitted with red data point; \textcolor{orange1}{Orange} \& \textcolor{red1}{Red} lines fitted without).}
\label{fig:polyfit}
\end{figure}

@@

Let $\mathcal{D}$ be a training data set with $n \in \mathbb{N}$ observations, where every data point consists of a target variable $y_i \in \mathbb{R}$ along with a number of corresponding inputs $x_i \in \mathbb{R}^p$. Given a linear function $f_\beta(x_i)$ of the inputs parametrized by coefficients $\beta \in \boldsymbol{\beta}$, $L(y_i f_\beta(x_i))$ is the \textit{Loss} function measuring the discrepancy between the actual target $y_i$ and the output of the linear function $f_\beta(x_i)$. According to \textcite{Vapnik1991}, the \textit{Empirical Risk Functional} is then 
\[
R_{emp}(\beta) = \frac{1}{n} \sum_{i=1}^n L(y_i, f_\beta(x_i)).
\]
\textit{Regularization} in this thesis' context refers then to adding some penalty function $J(\mathbf{\beta})$ dependent on the set of parameters $\mathbf{\beta}$, multiplied by some weighting parameter $\lambda$, to the empirical risk functional $R_{emp}(\mathbf{\beta})$. Thus, $R_{reg} = R_{emp}(\beta) + \lambda\cdot J(\mathbf{\beta})$. This results in the overall optimization problem
\begin{align}
&\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{R_{reg}(\beta)\right\}\nonumber\\
&=\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{R_{emp}(\beta) + \lambda\cdot J(\beta)\right\}\nonumber\\
&=\underset{\beta \in \boldsymbol{\beta}}{\argmin}\left\{\sum_{i=1}^n L(y_i, f_\beta(x_i)) + \lambda\cdot J(\beta)\right\}.\label{eq:genreg}
\end{align}

It is important to note here that $J(\mathbf{\beta})$ is only a function of the coefficients $\beta$ and neither the targets $y_i$ nor the inputs $x_i$. Therefore, it only affects the generalization error of the model, not the training error given by the empirical risk functional $R_{emp}(\mathbf{\beta})$.

@@

The concept of a penalized regression has been around for quite some time and been studied widely in various fields of scientific research. Arguably, this methodological approach to penalized regression started with the publication of two pieces of literature published by Arthur Hoerl and Robert Kennard in 1970 \parencite{HoerlKennard1970a, HoerlKennard1970b}. With these two papers, the authors introduced the widely known \textit{Ridge Regression}, which was derived from the previously known concept of Ridge analysis. At its core, the authors were trying to tackle the problem of high variances of the regression coefficients in high-dimensional problem settings. This shrinkage estimator, which uses the squared coefficient as a penalty term, "attempt[s] to shrink the coefficients to
reduce these variances, while adding some bias." \parencite{Hoerl2020} This closely resembles the previously discussed issue of the \textit{Bias-Variance-Tradeoff}, which has been discussed in the \textit{Under- and Overfitting} chapter in section \ref{sec:theorysuper} (Roger W. Hoerl, Arthur Hoerl's son, published a historical overview of the development of the concept of \textit{Ridge Regression} in 2020 \parencite{Hoerl2020}). The closed from solution of the Ridge estimator is given by
$$\hat{\mathbf{\beta}}_{Ridge} = \left(\mathbf{X}^T\mathbf{X}+\lambda \mathbf{I}\right)^{-1}\mathbf{X}^T\mathbf{y},$$
which adjusts the OLS estimator by shifting the main diagonal entries of the design matrix by $\lambda$ ($\lambda \geq 0$). It can be shown that this closed form estimator is equivalent to a Lagrangian problem of the following form \parencite{VanWieringen2015}:
$$\hat{\beta}_{Ridge}(\lambda) = \underset{\beta}{\argmin}\left\{\norm{\mathbf{y}-\mathbf{X}\mathbf{\beta}}_2^2 + \lambda\norm{\mathbf{\beta}}_2^2\right\}.$$
This resembles an example of the above defined regularization framework with squared residual loss and a penalty term of the form $\norm{\mathbf{\beta}}_2^2$, which only depends on the parameter $\beta$. In case of $\lambda$ being equal to zero, this reduces to the \textit{maximum likelihood (ML) estimator}.

@@

$$\norm{\mathbf{\beta}}_p =  \left(\sum_{i=1}^p|\beta_i|^p\right)^{1/p}.$$

Work published by researchers in the nineties, like the previously mentioned \textcite{FrankFriedman1993} or \textcite{Fu1998}, as well as more recent literature like \textcite{WangEtAl2020} have repeatedly shown there is no go-to-method to tackle regularization problems, as the effectiveness of a specific approach highly depends on the data situation at hand. Due to this particular situation in the literature, several other methods have been proposed in the recent years and decades. An approach combining the \textit{Ridge} and \textit{LASSO} methods is called \textit{Elastic Net} regularization and has been developed by \textcite{ZouHastie2005}. There, the authors elaborate on some of the shortcomings of the LASSO method. For example, in a special case where there are more predictors $p$ than data points $n$ ($p > n$), LASSO only selects up to $n$ variables due to the nature of the convex optimization problem. Should several of the included variables be highly \textit{pairwise} correlated with each other, \textit{LASSO} tends to only select one of these variables. At its core, \textit{Elastic Net Regularization} linearly combines the penalty terms of \textit{Ridge} and \textit{LASSO} regularization, yielding a loss function of the form:\\
$$J(\mathbf{\beta}) = \lambda_1 \norm{\mathbf{\beta}}_2^2 + \lambda_2 \norm{\mathbf{\beta}}_1.$$
The authors have	 shown that, especially when it comes to encouraging the aforementioned grouping effects, \textit{Elastic Net} tends to perform better than the \textit{LASSO}.\\

@@

Seeing this vast array of different pieces of literature immediately raises the question on advantages and disadvantages of specific methods and why so many have established itself in this particular field of study. Keeping in mind that the effectiveness of a method still highly depends on the data situation at hand, it is still important to distinguish methods based on its mathematical properties. \textcite{JohnVettamWu2022} have distinguished methods based on two broad criteria: (1) How does a penalty behave in small neighbourhoods around zero? (Is it smooth or singular at origin?) and (2) Is the underlying function convex or non-convex? To better incorporate the proposed concept of this thesis into this body of literature, I am proposing a third property to distinguish penalty terms: (3) How does the function behave at its limits? (Does it converge or diverge?) Such a classification tree can be found in figure \ref{fig:Classification}. \\

\begin{figure}
\begin{center}
\begin{tikzpicture}[
    level 1/.style={sibling distance=5.2cm},
    level 2/.style={sibling distance=2.6cm},
    level 3/.style={sibling distance=2.6cm, level distance=2.5cm},
    every node/.style={draw, rectangle, align=center}
]
\node {Classification of \\ Penalty Terms}
    child { node {Behaviour at Zero}
        child { node {Singularity} }
        child { node {Smooth at \\ Origin}}
    }
    child { node {(Non-)Convexity}
        child { node {Convex}}
        child { node {Non-Convex}}
    }
    child { node {Behaviour at Limits}
        child { node {Divergent}}
        child { node {Convergent}}
    };
\end{tikzpicture}  
\end{center}
\caption{A possible classification of regularization methods based on the mathematical properties of the penalty function} \label{fig:Classification}
\end{figure}
As already mentioned earlier in section \ref{sec:theorysuper}, a penalty function is a function dependent on the parameters of the model and it does not depend on the data at hand. Thus, changes in your data set does not effect the penalty directly, but only the overall optimization problem through the empirical risk functional $R_{emp}$. The following table \ref{tab:ClassificationOfPenalties} classifies some of the existing concepts based on the three criteria mentioned earlier. A visualisation of some of them can be found in figure \ref{fig:DifferentPenalties} in chapter \ref{subsec:comparepenalties}.\\

@@

LASSO 	& \textcite{Tibishirani1996}	& Singular	&	C 	& Div.\\
Ridge 	& \textcite{HoerlKennard1970a}	& Smooth		& C			& Div.\\
Gaussian & \textcite{JohnVettamWu2022}	& Smooth		& NC 	& Conv.\\
Ar(c)tan 	& \textcite{WangZhu2016}	& Singular 	& NC 	& Conv. \\
Elastic-Net & \textcite{ZouHastie2005} & Singular & C &\\
Bridge & \textcite{FrankFriedman1993}& Both &&\\
MCAP & \textcite{Zhang2010}& Singular & NC &\\
SCAD & \textcite{FanLi2001} & Singular & NC &\\
Laplace & \textcite{TrzaskoManduca2009} & Singular & NC &\\
Triple-Gamma & - & \makecell{Depending\\ on \\ parameters} & NC & Div. \\

@@

The main motivation for this thesis stems from a striking duality that exists between Bayesian shrinkage priors (discussed in chapter \ref{sec:BayesianShrinkagePriors}) and the frequentist approach using regularization. Both approaches aim at tackling the issue of model complexity and overfitting by making it necessary for the data to be more convincing that the value of an estimate is statistically significant different from zero. As mentioned, Bayesian statistics use specific prior distributions with usually a lot of mass around zero and heavy tails, whereas frequentist statisticians usually alter their optimization problems by incorporating penalty terms into into their loss functions. On a first glance, these two approaches have no immediate mathematical connection, however this turns out to be wrong.\\

@@

It turns out, as can be seen in \textcite[248--250]{ESL2009} and \textcite{VanWieringen2015} among others, that when choosing certain prior distributions and using standard assumptions in Bayesian modelling (ie the individual parameters of the parameter vector are independent a priori and the errors of the standard linear model are iid drawn from a normal distribution), that the mode of the posterior distributions correspond to the point estimates of a regularization approach in a frequentist setting. As this idea is essential to the entire thesis, a shortened derivation of this duality for the case of \textit{Ridge} regression will now be shown.\\

Assuming that a response variable $y_i$ follows a linear model with $p$ variables and errors $\epsilon_i \sim \mathcal{N}(0, \sigma^2)$, then the likelihood of the data is given by
$$\mathcal{L}(\mathbf{y}|\mathbf{X},\mathbf{\beta})=\frac{1}{(2\pi\sigma^2)^{n/2}}\exp\left(-\frac{1}{2\sigma^2}\sum_{i=1}^n(y_i-\mathbf{x_i}^T\mathbf{\beta})^2\right).$$

Under the previously mentioned assumption that $p(\beta) = \prod_{i=1}^p \beta_i$ and by choosing a Gaussian distribution with mean zero and variance $\tau^2$, we gain the prior distribution
$$p(\beta) = \frac{1}{(2\pi\tau^2)^{p/2}}\exp\left(-\frac{1}{2\tau^2}\norm{\mathbf{\beta}}_2^2\right).$$

Putting both together using Bayes' theorem and taking the $\log$ of the distribution yields
$$\log(p(\mathbf{\beta}|\mathbf{y},\mathbf{X})) \propto -\frac{1}{2\sigma^2}\sum_{i=1}^n (y_i-\mathbf{x_i}^T\mathbf{\beta})^2  - \frac{1}{2\tau^2}\norm{\mathbf{\beta}}_2^2.$$

Here, $\frac{1}{2\tau^2}$ can be viewed as the weighting parameter $\lambda$, which controls the strength of the regularization. The \textit{maximum-a-priori} estimate $\hat{\beta}_{MAP}$, which can be viewed as trying to find the $\underset{\beta}{\argmax}$ of the $\log$-posterior, can be used to construct the following optimization problem:

$$\hat{\beta}_{MAP} = \underset{\beta}{\argmin}\left\{\frac{1}{2\sigma^2}\sum_{i=1}^n \epsilon_i^2  + \frac{1}{2\tau^2}\norm{\mathbf{\beta}}_2^2\right\}.$$

This optimization problem constitutes the connection between the Bayesian and the frequentist approach as it can be interpreted as (1) finding the mode of the posterior distribution when using a Gaussian prior distribution and (2) finding the parameters which minimize the loss when using a Ridge regression with $\lambda = 1/(2\tau^2)$. The same connection exists for the case of \textit{LASSO} regression which can be derived using the same procedure and a double-exponential - also called \textit{Laplace} - distribution with mean zero and some scale parameter $\tau$ as the prior for the parameter vector $\mathbf{\beta}$.

@@

As already mentioned above, shrinkage priors are a Bayesian approach to dealing with statistical problems in high-dimensional settings. The prior is generally constructed for variance selection in sparse state space and TVP models and is named the Triple-Gamma-Prior as its representation is given by three gamma distributions. A key advantage of this prior is its unifying property. As can be seen in table 1 in \textcite{TGP2020}, special cases of the prior, depending on certain values in its hyperparameters, are the normal-exponential-Gamma, the horseshoe and the double gamma prior. In addition, the Bayesian LASSO can be seen as a special case of the triple-gamma-prior. Another advantage is its function as a bridge between spike-and-slab and continuous shrinkage prior as the hyperparameters can be used both for steering the shrinkage effect on noise while also preventing overshrinking the signal. Computationally, the shrinkage prior performs better than commonly used spike-and-slab priors as it is a continuous shrinkage prior which can more easily be implemented using Markov-Chain-Monte-Carlo (MCMC) methods.\\

@@

%----- Model Setup and Derivation
\newpage
\section{Model Setup and Derivation}\label{sec:modelderivation}
Coming to the theoretical framework of the \textit{triple-gamma-regularization}, let's assume we have a response variable $y$ and $p$ predictors along with $n$ data points. More formally, let $\mathbf{y}=[y_1  \quad y_2 \cdots y_n]^T$ and $\mathbf{x_i} = [x_{i1} \quad x_{i2} \cdots x_{ip}]^T$ with $\forall i\in \{1,...,n\}: y_i, x_{ij} \in \mathbb{R}$. Here, $x_i$ is the $i$-th data point, thus resulting in the design matrix $\mathbf{X} = [\mathbf{x_1} \quad \mathbf{x_2} \cdots \mathbf{x_n}]^T$.\\
